### 2019-09-16 ic-journal/multivariate.tex
Removed all reviewed highlights in the first part of the paper; re-enable highlights

In this section,  \hl{we discuss the formalisation of the} \multivariate
cases of Milner's ``unique solution of equations (for $\sim$)''
(Theorem~\ref{t:Mil89s1}) and our ``unique solution of rooted
contractions'' (Theorem~\ref{t:rcontraBisimulationU}).
\hl{This formalisation only supports finitely many
  equations/contractions and equation variables.\footnote{The original theorems
hold for even (uncountably) infinite equations and equation variables.}}
We chose these two theorems because they both depend on the same
concept of ``weakly guarded contexts'', and Lemma~\ref{lem:milner313}.
The work is an extension of the \univariate case~\cite{EPTCS276.10}
 in the sense that neither the CCS datatype nor
the SOS rules need to be changed.

The central problem of the \multivariate formalisation is the
representation of CCS equations (expressions, contexts).
\hl{In the} \univariate formalisation, $\lambda$-functions (of type
$CCS\rightarrow CCS$) are used to represent \univariate CCS
\hl{equations, and variable substitutions are simply 
calls to $\lambda$-functions on the target CCS terms.}
This idea cannot be extended to the \multivariate case, as
we do not have a fixed number of variables to deal with.

@@

% : dealing with another CCS-like type makes the
% whole formalisation work ugly and we loose the (possibly valuable) bisimilarity concepts and
% transitional semantics of CCS expressions; on the other hand,
% modifying the CCS datatype to just support the formalisation of
% unique-solution theorems is also not acceptable.

\hl{In this project, we have followed} Milner~\cite{Mil89}, using the
same alphabet for both agent variables and equation variables.
This has allowed a large reuse of the \univariate formalisation. \hl{For instance, 
the variable
substitution function used in SOS rules [\texttt{REC}] can be also
used for substituting eqaution variables.
However, a lot of care is needed in the proofs of many
fundamental lemmas, mostly due to the variable capture issues between
equation variables and agent variables.}

@@

We denote the set of all bound variables of a
given CCS term $E$ (we recall that these are variables that are bound in a recursion
subexpression of $E$) as $\bv{E}$ (or \HOLinline{\HOLConst{BV}\\\;\HOLFreeVar{E}} in HOL), and the set of  free
variables as $\fv{E}$ (or \HOLinline{\HOLConst{FV}\\\;\HOLFreeVar{E}} in HOL). Both \HOLinline{\HOLConst{BV}} and \HOLinline{\HOLConst{FV}} have the
type ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS} \HOLTokenTransEnd \ensuremath{\alpha} \HOLTokenTransEnd \HOLTyOp{bool}}'', i.e. functions taking CCS terms returning sets of
variables (of type \HOLinline{\ensuremath{\alpha}}).
% Milner didn't give their definitions explicitly. However, one can
% imagine that, for most CCS operators they just pass into
% subterms without collecting anything (and $\nil$ gives $\emptyset$).
For their definition the interesting cases are those of 
recursion and agent variables, shown below  (here \texttt{DELETE}
and \texttt{INSERT} are set-theoretic operators of HOL's
\texttt{pred_set} theory):
\begin{center}
\begin{tabular}{|l|l|}
\hline
\HOLConst{FV} \ensuremath{(}\HOLConst{var} \HOLFreeVar{X}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLTokenLeftbrace{}\HOLFreeVar{X}\HOLTokenRightbrace{} & \HOLConst{FV} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{p}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{FV} \HOLFreeVar{p} \HOLConst{DELETE} \HOLFreeVar{X} \\
\HOLConst{BV} \ensuremath{(}\HOLConst{var} \HOLFreeVar{X}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLSymConst{\HOLTokenEmpty{}} & \HOLConst{BV} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{p}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLFreeVar{X} \HOLConst{INSERT} \HOLConst{BV} \HOLFreeVar{p} \\
\hline
\end{tabular}
\end{center}
Furthermore, $E$ is a process, written \HOLinline{\HOLConst{IS_PROC} \HOLFreeVar{E}}, if
\hl{it does not contain any free variable}, i.e. $\fv{E} = \emptyset$:
\begin{alltt}
   \HOLConst{IS_PROC} \HOLFreeVar{E} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{FV} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\HOLTokenEmpty{}}\hfill{[IS_PROC_def]}
\end{alltt}
And a list of CCS processes can be asserted by \HOLinline{\HOLConst{ALL_PROC}} defined upon \HOLinline{\HOLConst{IS_PROC}}:
\begin{alltt}
   \HOLConst{ALL_PROC} \HOLFreeVar{Es} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{EVERY} \HOLConst{IS_PROC} \HOLFreeVar{Es}\hfill{[ALL_PROC_def]}
\end{alltt}

Note that, in CCS expression, the set of free and bound
variables need not be disjoint, e.g.\ in  $X + \recu X E$.
More importantly, when going
from a CCS expression to its sub-expressions, the set of free
variables may increase, while the set of bound variables \hl{either
remains the same or decreases}. For instance, $\fv{\recu X (\mu.X)} = \emptyset$, while
$\fv{\mu.X} = \{X\}$. This property of $\fv{\cdot}$ brings some
difficulities when doing proofs about CCS transitions. As an evidence,
we prove the following fundamental result about $\fv{\cdot}$:

@@

\hl{Milner said it can be proved by ``an easy action
induction''}~\cite{milner1990operational}.
From HOL's viewpoint, the so-called ``action induction'' is actually a higher-order \hl{application}

@@

rules (which essentially says \hl{that} \HOLinline{\HOLConst{TRANS}} is the smallest relation

@@

\hl{would have completed.}

@@

\hl{the subgoal for \texttt{SUM1} we need to prove}

@@

Knowing \HOLinline{\HOLConst{FV} \HOLFreeVar{E\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{FV} \HOLFreeVar{E}} we cannot prove \HOLinline{\HOLConst{FV} \HOLFreeVar{E\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{FV} \HOLFreeVar{E} \HOLConst{DELETE} \HOLFreeVar{X}}. However, if we knew \HOLinline{\HOLFreeVar{X} \HOLSymConst{\HOLTokenNotIn{}} \HOLConst{FV} \HOLFreeVar{E\sp{\prime}}}, then \HOLinline{\HOLConst{FV} \HOLFreeVar{E\sp{\prime}} \HOLConst{DELETE} \HOLFreeVar{X} \HOLSymConst{\ensuremath{=}} \HOLConst{FV} \HOLFreeVar{E\sp{\prime}}}, and \hl{then}

@@

\hl{would have completed.} Thus it remains to show that

@@

   \HOLConst{CCS_SUBST} \HOLFreeVar{fm} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}}
     \HOLKeyword{if} \HOLFreeVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLFreeVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLFreeVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLFreeVar{X}\ensuremath{)} \HOLFreeVar{E}\ensuremath{)}
     \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLFreeVar{fm} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_SUBST_rec]}
\end{alltt}
Note that the variable substitution only occurs on free
variables. In the case of $\recu X E$: if $X$ (a bound variable) is in
the domain of \HOLinline{\HOLFreeVar{fm}}, \HOLinline{\HOLConst{CCS_SUBST}} must continue the
substitution on $E$ using a reduced map~--- without $X$~--- so that all
occurences of $X$ in $E$ are correctly bypassed.
\hl{Since} \HOLinline{\HOLConst{CCS_SUBST}} only runs
once on each subterm, the possible free variables in  \HOLinline{\HOLFreeVar{Ps}} are never
 substituted. 
% Thus the process is indeed
% order-independent. 
Below we present some key lemmas about
\HOLinline{\HOLConst{CCS_SUBST}}, omitting the proofs:\footnote{Hereafter,
  some new logical constants from HOL's \texttt{pred_set}
and \texttt{list} theories are used (c.f.~\cite{holdesc} for more
details.): \HOLinline{\HOLConst{DISJOINT}} denotes set disjointness;
``\HOLinline{\HOLConst{set}\\\;\HOLFreeVar{Xs}}'' is the set converted from the list \HOLinline{\HOLFreeVar{Xs}};
\HOLinline{\HOLConst{ALL_DISTINCT}} says
all elements of a list are distinct with each other; \HOLinline{\HOLConst{MAP}} is
the mapping function from one list to another; \HOLinline{\HOLConst{EVERY}} means
each element of a list satisfies a predicate; \HOLinline{\HOLConst{ZIP}} creates a
new list from two lists of the same length, and each element of the new list is a pair of
elements, each from one of the old lists.}

@@

The next lemma plays an important role. It essentially swaps the order of two substitutions:
\begin{lemma}[\texttt{CCS_SUBST_nested}]
Under certain conditions \hl{(to get rid of substitution orders)}, two nested substitutions can be
converted into a single substitution where the targets are substituted first:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}}
   \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{C}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
   \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Es}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{C}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLConst{MAP} \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{Es}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{C}
\end{alltt}
\end{lemma}

@@

Recall in Section~\ref{ss:context}, \univariate contexts and their
guarded companions are defined as predicates over $\lambda$-expressions
of type $CCS\rightarrow CCS$. These predicates, such as
\HOLinline{\HOLConst{CONTEXT}}
(contexts/expressions), \HOLinline{\HOLConst{WG}} (weak guarded contexts),
\HOLinline{\HOLConst{SG}} (guarded contexts) and \HOLinline{\HOLConst{SEQ}} (sequential
contexts), are all defined inductively, i.e.~they are built from some ``holes''
in a bottom-up way. For instance, \HOLinline{\HOLConst{WG}\\\;\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\\\;\HOLSymConst{\ensuremath{\mid}}\\\;\HOLFreeVar{P}\ensuremath{)}}
holds because \HOLinline{\HOLConst{WG}\\\;\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}}
holds \hl{as a base case of the inductive definition of} \HOLinline{\HOLConst{WG}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{a}. \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}\hfill{[WG1]}
\end{alltt}
\hl{For} any agent variable $X$, we have \hl{by $\beta$-reduction:}
\HOLinline{\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{P}\ensuremath{)}\\\;\ensuremath{(}\HOLConst{var}\\\;\HOLFreeVar{X}\ensuremath{)}\\\;\HOLSymConst{\ensuremath{=}}\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{P}} (or ``$l.X + P$'' in textbook
notation) is weakly guarded (c.f.~Def.~\ref{def:guardness}).

>>>>

In this section,  we discuss the formalisation of the \multivariate
cases of Milner's ``unique solution of equations (for $\sim$)''
(Theorem~\ref{t:Mil89s1}) and our ``unique solution of rooted
contractions'' (Theorem~\ref{t:rcontraBisimulationU}).
\hl{This formalisation only supports finitely many
  equations/contractions and equation variables.}\footnote{\hl{The original theorems
hold for even (uncountably) infinite equations and equation variables.}}
We chose these two theorems because they both depend on the same
concept of ``weakly guarded contexts'', and Lemma~\ref{lem:milner313}.
The work is an extension of the \univariate case~\cite{EPTCS276.10}
 in the sense that neither the CCS datatype nor
the SOS rules need to be changed.

The central problem of the \multivariate formalisation is the
representation of CCS equations (expressions, contexts).
In the \univariate formalisation, $\lambda$-functions (of type
$CCS\rightarrow CCS$) are used to represent \univariate CCS
equations, and variable substitutions are simply 
calls to $\lambda$-functions on the target CCS terms.
This idea cannot be extended to the \multivariate case, as
we do not have a fixed number of variables to deal with.

@@

In this project, we have followed Milner~\cite{Mil89}, using the
same alphabet for both agent variables and equation variables.
This has allowed a large reuse of the \univariate formalisation.
For instance, the variable
substitution function used in SOS rules \texttt{REC} can be also
used for substituting eqaution variables.
However, a lot of care is needed in the proofs of many
fundamental lemmas, mostly due to the variable capture issues between
equation variables and agent variables.

@@

We denote the set of all bound variables of a
given CCS term $E$ (we recall that these are variables that are bound in a recursion
subexpression of $E$) as $\bv{E}$ (or \HOLinline{\HOLConst{BV}\\\;\HOLFreeVar{E}} in HOL), and the set of  free
variables as $\fv{E}$ (or \HOLinline{\HOLConst{FV}\\\;\HOLFreeVar{E}} in HOL). Both \HOLinline{\HOLConst{BV}} and \HOLinline{\HOLConst{FV}} have the
type ``\HOLinline{\ensuremath{(}\ensuremath{\alpha}, \ensuremath{\beta}\ensuremath{)} \HOLTyOp{CCS} \HOLTokenTransEnd \ensuremath{\alpha} \HOLTokenTransEnd \HOLTyOp{bool}}'', i.e. functions taking CCS terms returning sets of
variables (of type \HOLinline{\ensuremath{\alpha}}).
% Milner didn't give their definitions explicitly. However, one can
% imagine that, for most CCS operators they just pass into
% subterms without collecting anything (and $\nil$ gives $\emptyset$).
For their definition the interesting cases are those of 
recursion and agent variables, shown below  (here \texttt{DELETE}
and \texttt{INSERT} are set-theoretic operators of HOL's
\texttt{pred_set} theory):
\begin{center}
\begin{tabular}{|l|l|}
\hline
\HOLConst{FV} \ensuremath{(}\HOLConst{var} \HOLFreeVar{X}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLTokenLeftbrace{}\HOLFreeVar{X}\HOLTokenRightbrace{} & \HOLConst{FV} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{p}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{FV} \HOLFreeVar{p} \HOLConst{DELETE} \HOLFreeVar{X} \\
\HOLConst{BV} \ensuremath{(}\HOLConst{var} \HOLFreeVar{X}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLSymConst{\HOLTokenEmpty{}} & \HOLConst{BV} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{p}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLFreeVar{X} \HOLConst{INSERT} \HOLConst{BV} \HOLFreeVar{p} \\
\hline
\end{tabular}
\end{center}
Furthermore, $E$ is a process, written \HOLinline{\HOLConst{IS_PROC} \HOLFreeVar{E}}, if
it does not contain any free variable, i.e. $\fv{E} = \emptyset$:
\begin{alltt}
   \HOLConst{IS_PROC} \HOLFreeVar{E} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{FV} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLSymConst{\HOLTokenEmpty{}}\hfill{[IS_PROC_def]}
\end{alltt}
And a list of CCS processes can be asserted by \HOLinline{\HOLConst{ALL_PROC}} defined upon \HOLinline{\HOLConst{IS_PROC}}:
\begin{alltt}
   \HOLConst{ALL_PROC} \HOLFreeVar{Es} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{EVERY} \HOLConst{IS_PROC} \HOLFreeVar{Es}\hfill{[ALL_PROC_def]}
\end{alltt}

Note that, in CCS expression, the set of free and bound
variables need not be disjoint, e.g.\ in  $X + \recu X E$.
More importantly, when going
from a CCS expression to its sub-expressions, the set of free
variables may increase, while the set of bound variables either
remains the same or decreases. For instance, $\fv{\recu X (\mu.X)} = \emptyset$, while
$\fv{\mu.X} = \{X\}$. This property of $\fv{\cdot}$ brings some
difficulities when doing proofs about CCS transitions. As an evidence,
we prove the following fundamental result about $\fv{\cdot}$\hl{\mbox{~\citep[p.~1209]{milner1990operational}}}:

@@

\hl{Milner said it can be proven by ``an easy action induction''.}
From HOL's viewpoint, the so-called ``action induction'' is actually a higher-order application

@@

rules (which essentially says that \HOLinline{\HOLConst{TRANS}} is the smallest relation

@@

would have completed.

@@

the subgoal for \texttt{SUM1} we need to prove

@@

Knowing \HOLinline{\HOLConst{FV} \HOLFreeVar{E\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{FV} \HOLFreeVar{E}} we cannot prove \HOLinline{\HOLConst{FV} \HOLFreeVar{E\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{FV} \HOLFreeVar{E} \HOLConst{DELETE} \HOLFreeVar{X}}. However, if we knew \HOLinline{\HOLFreeVar{X} \HOLSymConst{\HOLTokenNotIn{}} \HOLConst{FV} \HOLFreeVar{E\sp{\prime}}}, then \HOLinline{\HOLConst{FV} \HOLFreeVar{E\sp{\prime}} \HOLConst{DELETE} \HOLFreeVar{X} \HOLSymConst{\ensuremath{=}} \HOLConst{FV} \HOLFreeVar{E\sp{\prime}}}, and then

@@

would have completed. Thus it remains to show that

@@

   \HOLConst{CCS_SUBST} \HOLFreeVar{fm} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}}
     \HOLKeyword{if} \HOLFreeVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLFreeVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLFreeVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLFreeVar{X}\ensuremath{)} \HOLFreeVar{E}\ensuremath{)}
     \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLFreeVar{fm} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_SUBST_rec]}
\end{alltt}
Note that the variable substitution only occurs on free
variables. In the case of $\recu X E$: if $X$ (a bound variable) is in
the domain of \HOLinline{\HOLFreeVar{fm}}, \HOLinline{\HOLConst{CCS_SUBST}} must continue the
substitution on $E$ using a reduced map~--- without $X$~--- so that all
occurences of $X$ in $E$ are correctly bypassed.
Since \HOLinline{\HOLConst{CCS_SUBST}} only runs
once on each subterm, the possible free variables in  \HOLinline{\HOLFreeVar{Ps}} are never
 substituted. 
% Thus the process is indeed
% order-independent. 
Below we present some key lemmas about
\HOLinline{\HOLConst{CCS_SUBST}}, omitting the proofs:\footnote{Hereafter,
  some new logical constants from HOL's \texttt{pred_set}
and \texttt{list} theories are used (c.f.~\cite{holdesc} for more
details.): \HOLinline{\HOLConst{DISJOINT}} denotes set disjointness;
``\HOLinline{\HOLConst{set}\\\;\HOLFreeVar{Xs}}'' is the set converted from the list \HOLinline{\HOLFreeVar{Xs}};
\HOLinline{\HOLConst{ALL_DISTINCT}} says
all elements of a list are distinct with each other; \HOLinline{\HOLConst{MAP}} is
the mapping function from one list to another; \HOLinline{\HOLConst{EVERY}} means
each element of a list satisfies a predicate; \HOLinline{\HOLConst{ZIP}} creates a
new list from two lists of the same length, and each element of the new list is a pair of
elements, each from one of the old lists.}

@@

The next lemma plays an important role. It essentially swaps the order of two substitutions:
\begin{lemma}[\texttt{CCS_SUBST_nested}]
Under certain conditions (to get rid of substitution orders), two nested substitutions can be
converted into a single substitution where the targets are substituted first:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}}
   \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{C}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
   \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Es}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{C}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLConst{MAP} \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{Es}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{C}
\end{alltt}
\end{lemma}

@@

Recall in Section~\ref{ss:context}, \univariate contexts and their
guarded companions are defined as predicates over $\lambda$-expressions
of type $CCS\rightarrow CCS$. These predicates, such as
\HOLinline{\HOLConst{CONTEXT}}
(contexts/expressions), \HOLinline{\HOLConst{WG}} (weak guarded contexts),
\HOLinline{\HOLConst{SG}} (guarded contexts) and \HOLinline{\HOLConst{SEQ}} (sequential
contexts), are all defined inductively, i.e.~they are built from some ``holes''
in a bottom-up way. For instance, \HOLinline{\HOLConst{WG}\\\;\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\\\;\HOLSymConst{\ensuremath{\mid}}\\\;\HOLFreeVar{P}\ensuremath{)}}
holds because \HOLinline{\HOLConst{WG}\\\;\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}}
holds as a base case of the inductive definition of \HOLinline{\HOLConst{WG}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{a}. \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLBoundVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}\hfill{[WG1]}
\end{alltt}
For any agent variable $X$, we have by $\beta$-reduction:
\HOLinline{\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{P}\ensuremath{)}\\\;\ensuremath{(}\HOLConst{var}\\\;\HOLFreeVar{X}\ensuremath{)}\\\;\HOLSymConst{\ensuremath{=}}\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{P}} (or ``$l.X + P$'' in textbook
notation) is weakly guarded (c.f.~Def.~\ref{def:guardness}).

It is possible to inductively assert \multivariate contexts and their
guarded variants, but care is needed as  we need to exclude the cases where
free variables occur inside a recusion operator. A more elegant
 solution is to define \multivariate contexts (and variants) upon the
existing \univariate definitions, which is the approach we have followed.
For example, to see the weak guardedness of $a.X + b.X + c.Y$, it is
natural to focus on each (free) variable: if we substitute each
occurrence of just one variable, say $X$, with a ``hole'' and see the
resulting term as a $\lambda$-function, then all such $\lambda$-functions
must satisfy \HOLinline{\HOLConst{WG}}, i.e.~the following results hold:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{Y}\ensuremath{)}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}
\end{alltt}
Also note that \HOLinline{\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\\\;\HOLSymConst{\ensuremath{=}}\\\;\ensuremath{[}\HOLFreeVar{t}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\\\;\ensuremath{(}\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\ensuremath{)}}
can be expressed as a \univariate substitution of the original term.
This idea leads to the following definitions:
\begin{alltt}
   \HOLConst{context} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{CONTEXT} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}\hfill{[context_def]}
